Wireless service accounting based on Ethereum state channels - Master Thesis

### 2019-10-06 Paper.tex
Add 5G lore, start software description

Since spectrum is a scarce resource, it can
be tokenized being suited this
way for blockchain/DLT technologies business models;
in addition, since
atmospheric spectrum can be considered a public
asset, it's managed in last terms by nations' governments
and it's licensing and management should feature
a high level of transparency and at the same time
allow for efficiency and fair-competition making it
feasible to model on public blockchain systems.

@@

\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 2. Raspberry Pi 3 Model B.
\\
\end{center}

@@

In the recent years blockchain technologies
and in particular smart contracts have made possible
an innovative way to rule business and
workflows between organizations.

@@

In the case of 5G technologies and due to the
requirement that different operators shall be able
to dynamically share the same resources (i.e. neutral host
infrastructure). At the same time in order to manage
service operation efficiently through its whole
life-cycle there's the need to automate as much as
possible all the operations; this requirements
perfectly fit smart contract capabilities, in order
to model and operate over a defined set of rules
supporting payments, escrows, rewards or penalties
depending on different conditions.

\end{multicols}
\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g-y.eps}
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g2-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 3. Anonymicity, identity, KyC and GDPR.
\\
\end{center}
\begin{multicols}{2}

\end{multicols}
\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 4. Slicing model.
\\
\end{center}
\begin{multicols}{2}

\end{multicols}
\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc1-y.eps}
  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc2-y.eps}
  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc3-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 5. State channels support instant portability
between different operators with minimal fee.
\\
\end{center}
\begin{multicols}{2}

@@

\section{Quality considerations}\label{sec:proofs}
Para validar la PoC, se ha definido una batería de tests. Por ejemplo, dado que los smart contract normalmente manejan dinero, es esencial asegurarse de que su número de fallos y vulnerabilidades sea bajo [Hegedus 18]. Para ayudar a los desarrolladores y hacer más madura la tecnología, necesitamos herramientas de análisis [ConsenSys 19].
\subsection{Automated tests} \label{ch:tests}
(Por completar)

\section{Deployment}\label{sec:deploy}
Docker es una herramienta que permite desplegar aplicaciones dentro de contenedores de software. Esto puede ser útil para la Raspberry Pi porque permite a los usuarios ejecutar aplicaciones con muy poca sobrecarga, siempre y cuando la aplicación esté empaquetada dentro de una imagen Docker. Simplemente instalamos Docker y ejecutamos el contenedor sobre Raspbian/ARM. Éste despliega el proyecto realizado.

La secuencia de comandos a ejecutar es:
\\
\begin{flushleft}
\$ sudo apt-get install git
\\
\$ git clone https://github.com/ethereum-internet-access/docker.git
\\
\$ cd docker
\\
\$ sh ./install-project-inside-docker-container.sh
\end{flushleft}
\section{Viabilidad de negocio}\label{sec:business}
(Por completar)
\subsection{Legislación} \label{ch:legislation}
(Por completar)
\section{Conclusiones}\label{sec:conclusions}
El mundo en el que vivimos es interconectado.
En él Internet juega un papel fundamental.
Se ha contribuido a la construcción de
una PoC. Ésta se ha basado en la medida de
lo posible en hardware libre
(Raspberry Pi) ya que goza de buena popularidad
para el desarrollo de prototipos.
Como debilidad del pago con ETH, se puede
achacar la volatilidad de su precio (véase figura 3).
Esto conlleva variaciones en
las comisiones (\textit{fees}) de una transacción Ethereum.
\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase.eps}
%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
%\end{figure}
\\
Figura 3. Variación cotización ETH (11/08/2019). Fuente: [Coinbase 19]
\\
\end{center}
(Por completar)
\section{Líneas futuras}\label{sec:future}
Como líneas futuras y complemento a este trabajo, con más tiempo, se plantea:
\subsection{Añadir forma de pago DAI} \label{ch:dai}
DAI es una criptomoneda estable (\textit{stablecoin}) que funciona con ETH y que intenta mantener un valor de 1\$. A diferencia de los billetes de banco centralizados, DAI no está respaldado por dólares estadounidenses en una cuenta bancaria. En cambio, está respaldado por garantías en la plataforma Maker [MakerDAO 19].
\subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
Anunciado en Junio de 2019 y a la venta. Es un hardware que ofrece mejores prestaciones. Por ejemplo, el procesador es un Broadcom BCM2711. Un ARM Cortex-A72 de cuatro núcleos y a 1,5 GHz. Hasta tres veces más eficiente (\textit{benchmark}) que el modelo anterior (Raspberry Pi 3 Modelo B+). También da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
\subsection{Usar Docker Swarm} \label{ch:dockerswarm}
Se podría presentar una aplicación descentralizada basada en microservicios. Por ejemplo en microservicios, un contenedor Docker (como en la sección 4) podría verse como un servicio. Entre las ventajas que aporta microservicios estarían:
\begin{itemize}
 \item Pequeños
 \item Independientes
 \item Despliegue sencillo
 \item Reutilizables
 \item Externalización
 \item Escalabilidad
 \newline
\end{itemize}
Para ello, Docker ofrece un orquestador (de código abierto) llamado Docker Swarm (véase figura 4). Un nodo \textit{worker} sería nuestra Raspberry Pi y un nodo \textit{manager} sería un centro de control.
\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker.eps}
%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
%\end{figure}
\\
Figura 4. Diagrama Docker Swarm. Fuente: [Docker 19]
\\
\end{center}
\section{Repositorio}\label{sec:repository}
El código fuente generado en el proyecto está publicado en un repositorio público de GitHub bajo licencia MIT.
\newline\newline
\url{https://github.com/ethereum-internet-access}
\newline
\begin{thebibliography}{9}
\bibitem{spectrumnh}
  \href{https://uk5g.org/media/uploads/resource_files/Spectrum\_NH\_discussion\_paper\_20Feb19.pdf}{5G Spectrum and Neutral Hosting}, Discussion paper, February 2019,
  DCMS Phase 1 5G Testbeds \& Trials Programme
\bibitem{5g} \href{https://en.wikipedia.org/wiki/5G}{https://en.wikipedia.org/wiki/5G}
\bibitem{sdns} Ian F. Akyildiz and Shih-Chun Lin and Pu Wang,
  \textit{Wireless software-defined networks (W-SDNs) and network function virtualization (NFV) for 5G cellular systems:
    An overview and qualitative evaluation}, Computer Networks, 93, 2015
\bibitem{nfv} Jose Ordonez-Lucena et al.
  \textit{Network Slicing for 5G with SDN/NFV: Concepts, Architectures, and Challenges},
   IEEE Communications Magazine ( Volume: 55 , Issue: 5 , May 2017 )
\bibitem{fon} \href{https://es.wikipedia.org/wiki/La_Fonera}{https://es.wikipedia.org/wiki/La\_Fonera}
\bibitem{hotspotme} \href{https://devpost.com/software/hotspot-me}{https://devpost.com/software/hotspot-me}
\bibitem{hostapd} \href{https://w1.fi/hostapd/}{https://w1.fi/hostapd/}
\bibitem{iptables} \href{https://linux.die.net/man/8/iptables}{https://linux.die.net/man/8/iptables}
\bibitem{dnsmasq} \href{https://en.wikipedia.org/wiki/Dnsmasq}{https://en.wikipedia.org/wiki/Dnsmasq}
\bibitem{osmc} \href{https://osmc.tv/}{https://osmc.tv/}

>>>>

Since spectrum is a scarce resource, it can
be tokenized being suited this
way for blockchain/DLT technologies business models;
in addition, since
atmospheric spectrum can be considered a public
asset, it's managed at the end by nations' governments
and its licensing and management should feature
a high level of transparency and at the same time
allow for efficiency and fair-competition making it
feasible to model on public blockchain systems.

@@

\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 4. Raspberry Pi 3 Model B.
\\
\end{center}

\subsection{Detailed software description}

This section includes a detailed rationale software description of all
the components and subcomponents.

\section{Quality considerations}\label{sec:proofs}

To validate the PoC, a battery of tests has been defined. For example, given
that smart contract normally handle money, it is essential to ensure that
its number of failures and vulnerabilities is low [Hegedus 18].
To help developers and make technology more mature,
we need analysis tools [ConsenSys 19].

For this project, we used Mythril from ConsenSys. Mythril is a
security analysis tool for EVM (Ethereum Virtual Machine) bytecode.
It detects security vulnerabilities in smart contracts built for Ethereum.

We analysed the Solidity-based smart contract called
StateChannel.sol (of this project).

The installation steps are as follows in:

https://github.com/ethereum-internet-access/mythril
Firstly, we analysed the local source code:

\begin{verbatim}
$ cd contracts
$ myth analyze StateChannel.sol \
  > ./security-report
$ cat ./security-report
\end{verbatim}

Secondly, we analysed on-chain smart contract.
Because the smart contract was deployed on Ropsten testnet we executed:

\begin{verbatim}
$ myth analyze --rpc \
  infura-ropsten -a \
  contract_address \
  > ./security-report
$ cat ./security-report
\end{verbatim}

\section{Deployment}\label{sec:deploy}

Docker is a tool that allows you to deploy applications
inside of software containers. This can be useful for our
Raspberry Pi because it allows users to run applications with
very little overhead, as long as the application
is packaged inside of a Docker image.

We simply install Docker and run the container over Raspbian/ARM.
It would deploy the done project. The script to execute is as follows:

\begin{verbatim}
$ sudo apt-get update
$ sudo apt-get upgrade
$ sudo apt-get install curl
$ curl -sSL https://get.docker.com | sh
$ sudo apt-get install git
$ git clone \
  https://github.com/ \
  ethereum-internet-access/docker.git
$ cd docker
$ sh ./install-project\
  -inside-docker-container.sh
\end{verbatim}

@@

\subsection{5G Technologies and scarcity model}
\vspace{0.35cm}

In the recent years blockchain technologies
and in particular smart contracts have made possible
an innovative way to rule business and
processes between organizations.

@@

In the case of 5G technologies for practical reasons
and the high density of antennas required,
different operators shall be able
to dynamically share the same resources (i.e. neutral host
infrastructure). At the same time in order to manage
service operation efficiently through its whole
life-cycle there's the need to automate as much as
possible all the operations; this requirements
perfectly fit smart contract capabilities, in order
to model and operate over a defined set of rules
supporting payments, escrows, rewards or penalties
depending on different conditions.

\end{multicols}
\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 5. Slicing model.
\\
\end{center}
\begin{multicols}{2}

As an example, in figure 5 it's shown how three virtual operators over
different infrastructures shared or not in a slice-model of a neutral host
infrastructure; a slice can be described as a dynamic virtualization
allowing multitenancy of some physical resources like access points,
physical switches and datacenters including mobile edge computing
appliances; this is the case of currently in execution projects like
5GCity\cite{5gcity} where a distributed cloud
and radio platform for municipalities and infrastructure owners acting
as 5G neutral hosts is being developed and tested. Platforms
targeting 5G business models such as 5GCity can benefit from adding
this smart contract business logic in order to bill end-users and automate
all the accounting part.

\subsection{Know your customer issue}

Although this is a promising approach; there's an issue with it regarding
regulation, in particular in the European Union, there's the requirement
that a telecommunication service provider must know their customers; this
is required for example to allow judge requirements of users identity
in case there's some evidence of illegal activities. As a single payment
system, state channel smart contracts are pseudoanymous hence do not
support the \textit{Know Your Customer - KyC} required by regulation.

To take into account the problem illustrated in previous paragraph
we can take as a basis a model of self-sovereign identity\cite{allen}
and in particular the one recently anounced by the authorities
of Catalonia government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that the credential can be linked
to some administration database that could be queried only by a judge
requirement and guarantee the users identities are recognized

\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g-y.eps}
  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g2-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 6. Upper figure shows a scenario where different users connect
to antennas controlled by a state channel smart contract like the one
described in this thesis; users remain pseudoanonymous and hence, the
schema is not regulation compliant regarding the KyC that a
telecommunications
operator should guarantee in case of legal issues. On the other hand,
relying on a self-sovereign decentralized identity bottom figure shows
up how the citizens can authenticate themselves to the system with the
guarantee there's a public administration validating their identities.
\\
\end{center}

\begin{multicols}{2}
by a public administration; this way, the service provider (\textit{in
  our case a telecommunications operator}) does not need to know
anything about a customer appart from a public key and a valid
credential signed by the administration. This schema could be a real game
changer regarding operators competitivity since they get rid-off
from several and costly processes that do not add any value to the service
they provide:

\begin{itemize}
\item Fully automated accounting and billing layer.
\item Regulation compliance without KyC on-boarding process.
\item GDP compliance by default. Operator does not need to know
  user's name, bank account, etc.
\end{itemize}

\subsection{Instant portability}

This approach to the way customers interact with telecommunication
operators relying on a publicly recognized self-sovereign identity
can support a new way of handling portability granting full customer
empowerment when they want to switch from one operator to another;
in figure 5, is represented a process on a user opens a channel
in operator's A smart-contract; after that begins transacting with
this first operator paying as they consume the access service
sending micro-payments to the operator.

After that, user decides to switch to a second operator, opening a channel
over operator B's contract; since operator B, stops receiving the signed
micropayments, opts for closing the channel backdrawing the earnings in operator's
account and return value for the user. Such kind of process can be accomplisehd
in a matter of seconds for a few cents fee literally obliterating any
operator lock-in mechanism thus supporting a really zero friction service market.

\begin{center}
%\begin{figure}[h]
  %  \centering
  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc1-y.eps}
  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc2-y.eps}
  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc3-y.eps}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figure 5. State channels support instant portability
between different operators with minimal fee.
\\
\end{center}
\begin{multicols}{2}

  \subsection{Towards telco market full tokenization}

  This model can be extended without limits to the full wireless
  telecommunication market tokenization; as depicted in figure 6,
  from the spectrum auctioning, including infrastructure providers
  virtualizing their appliances for virtual operators, end to users
  service provision and accounting relying on a self-sovereign
  identity system.

@@

%% \section{Viabilidad de negocio}\label{sec:business}
%% (Por completar)
%% Comento el tema viabilidad de negocio porque ya está discutida
%% en Further steps

\subsection{Regulation considerations} \label{ch:regulation}

TODO; talk about GDPR, telco KyC, european market regulation, etc.

\section{Conclusions}\label{sec:conclusions}

World we live today is interconnected and wireless
internet access has a foundamental role. In this project
a Proof of Concept has been build. This PoC based in
commodity free hardware asset (Raspberry Pi) can serve
as an example of end-user wireless provision pay-as-you-go
using a state channel Solidity smart contract; KyC issues
and identity consideration have been discussed also leading
to the conclusion that complementing this approach with
what a self-sovereign identity can provide there's no need
to a telecommunication operator to perform a KyC process
in order to provision a service to a client.

One of the major problems with this approach has to do
with ETH-fiat exchange volatility
(\textit{see figure 7}), anyway, including
a feature to regulate minute price in the state channel
contract and taking into account gas prices can adjust
in order to properly
represent computation, bandwidth and storage costs this
problem can be completely avoided.

\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase.eps}
%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
%\end{figure}
\\
Figure 7. ETH Price chart (11/08/2019). Source: [Coinbase 19]
\\
\end{center}

%% \subsection{Añadir forma de pago DAI} \label{ch:dai}
%% DAI es una criptomoneda estable (\textit{stablecoin}) que funciona con ETH y que intenta mantener un valor de 1\$. A diferencia de los billetes de banco centralizados, DAI no está respaldado por dólares estadounidenses en una cuenta bancaria. En cambio, está respaldado por garantías en la plataforma Maker [MakerDAO 19].
%% \subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
%% Anunciado en Junio de 2019 y a la venta. Es un hardware que ofrece mejores prestaciones. Por ejemplo, el procesador es un Broadcom BCM2711. Un ARM Cortex-A72 de cuatro núcleos y a 1,5 GHz. Hasta tres veces más eficiente (\textit{benchmark}) que el modelo anterior (Raspberry Pi 3 Modelo B+). También da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
\subsection{Usar Docker Swarm} \label{ch:dockerswarm}
Se podría presentar una aplicación descentralizada basada en microservicios. Por ejemplo en microservicios, un contenedor Docker (como en la sección 4) podría verse como un servicio. Entre las ventajas que aporta microservicios estarían:
\begin{itemize}
 \item Pequeños
 \item Independientes
 \item Despliegue sencillo
 \item Reutilizables
 \item Externalización
 \item Escalabilidad
 \newline
\end{itemize}
Para ello, Docker ofrece un orquestador (de código abierto) llamado Docker Swarm (véase figura 4). Un nodo \textit{worker} sería nuestra Raspberry Pi y un nodo \textit{manager} sería un centro de control.
\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker.eps}
%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
%\end{figure}
\\
Figura 4. Diagrama Docker Swarm. Fuente: [Docker 19]
\\
\end{center}

\section{Repositorio}\label{sec:repository}
El código fuente generado en el proyecto está publicado en un repositorio público de GitHub bajo licencia MIT.
\newline\newline
\url{https://github.com/ethereum-internet-access}
\newline
\begin{thebibliography}{9}
\bibitem{spectrumnh}
  \href{https://uk5g.org/media/uploads/resource_files/Spectrum\_NH\_discussion\_paper\_20Feb19.pdf}{5G Spectrum and Neutral Hosting}, Discussion paper, February 2019,
  DCMS Phase 1 5G Testbeds \& Trials Programme
\bibitem{5g} \href{https://en.wikipedia.org/wiki/5G}{https://en.wikipedia.org/wiki/5G}
\bibitem{sdns} Ian F. Akyildiz and Shih-Chun Lin and Pu Wang,
  \textit{Wireless software-defined networks (W-SDNs) and network function virtualization (NFV) for 5G cellular systems:
    An overview and qualitative evaluation}, Computer Networks, 93, 2015
\bibitem{nfv} Jose Ordonez-Lucena et al.
  \textit{Network Slicing for 5G with SDN/NFV: Concepts, Architectures, and Challenges},
   IEEE Communications Magazine ( Volume: 55 , Issue: 5 , May 2017 )
\bibitem{fon} \href{https://es.wikipedia.org/wiki/La_Fonera}{https://es.wikipedia.org/wiki/La\_Fonera}
\bibitem{hotspotme} \href{https://devpost.com/software/hotspot-me}{https://devpost.com/software/hotspot-me}
\bibitem{hostapd} \href{https://w1.fi/hostapd/}{https://w1.fi/hostapd/}
\bibitem{iptables} \href{https://linux.die.net/man/8/iptables}{https://linux.die.net/man/8/iptables}
\bibitem{dnsmasq} \href{https://en.wikipedia.org/wiki/Dnsmasq}{https://en.wikipedia.org/wiki/Dnsmasq}
\bibitem{osmc} \href{https://osmc.tv/}{https://osmc.tv/}
\bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}
\bibitem{allen}  \href{Christopher Allen, The path to self sovereign identity, 2016}
  {http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
\bibitem{identicat}
  \href{Catalonia to Build DLT-Based Identity, Coindesk, 2019}{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}
